---
title: "waterfowl"
author: "Tati Micheletti"
date: "17 February 2021"
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hold")
```

# Overview

This module forecasts waterfowl habitat suitability index for up to 17 species of waterfowl for a given study area within the Western Boreal Region.

# Usage

The following inputs are expected (but do present defaults if not provided):  

*speciesURL:*  Data.table with 2 columns, species (4-letter code) and URL (url of google drive FOLDER where the fitted models from biomod2 were saved). It is important that the structure of the models is kept intact as the predict function depends on it.  
*species:* Character vector with species (4-letter code). These need to be available in speciesURL table.  
*studyArea:* Study area for the prediction. Defaults to the NWT.
*rasterToMatch:* All spatial outputs will be reprojected and resampled to it (i.e. projection and resolution).
*climateURLTable:* Data.table with information on decade, url and climate model.

studyAreaWBI_url <- "https://drive.google.com/file/d/19agsb9pzqsS_ZW7aS4MHa4QUiEp3R4fj/view?usp=sharing"

```{r setup}
library("Require")
Require("SpaDES")
Require("data.table")
Require("googledrive")
Require("raster")
Require("biomod2")

wd <- getwd() # This will be the waterfowl folder

rasterOptions(tmpdir = checkPath(file.path(wd, "rasterTemp"), create = TRUE))

setPaths(cachePath = checkPath(file.path(wd, "cache"), create = TRUE), 
         inputPath = checkPath(file.path(wd, "inputs"), create = TRUE),
         outputPath = checkPath(file.path(wd, "outputs"), create = TRUE))

```

If you want to run future predictions, run the following code:

```{r module_usage}

times <- list(start = 2050, end = 2070)

modules <- list("waterfowl")

objects <- list()
inputs <- list()
outputs <- list()

mySim <- simInit(times = times, params = parameters, modules = modules,
                 objects = objects)

mySimOut <- spades(mySim)
```

It is also possible to assess this model for current time (i.e. 2011), albeit outside of the forecasting SpaDES module:

```{r run2011}

speciesURL <- data.table(species = c("REDH", "BWTE", "BUFF"),
                         URL = c("https://drive.google.com/file/d/1Sw8V93gv3eJNTuCeD3ZD6jRScVR5h-Yb/view?usp=sharing",
                                 "https://drive.google.com/file/d/1G7Zq5BBgQtoxBgiGJmojbbIZpb8vZY9I/view?usp=sharing",
                                 "https://drive.google.com/file/d/1C-oTdV9yzS6OEiO_fan7uoVACL-d1ajD/view?usp=sharing"))

covariates <- "https://drive.google.com/file/d/1UXl5zBG9D39G9FpJeCYbD3RU6WA6LNq9/view?usp=sharing"

# Download the models and data for each species
  speciesPaths <- rbindlist(lapply(1:NROW(speciesURL), function(index){
    sp <- speciesURL[index, species]
    URL <- speciesURL[index, URL]
    if (!dir.exists(file.path(Paths$inputPath, 
                   sp))){
      filePath <- file.path(Paths$inputPath,
                            paste0(sp, ".zip"))
      
      drive_download(file = URL, path = filePath)
      unzip(zipfile = filePath, exdir = Paths$inputPath)
      speciesPath <- data.table(species = sp,
                                spInputPath = file.path(Paths$inputPath, 
                                                        sp))
      } else {
      speciesPath <- data.table(species = sp,
                                spInputPath = file.path(Paths$inputPath, 
                                                        sp))
    } 
    return(speciesPath)
  }))

# Retrieve current environmental covariate layers
allCov <- prepInputs(targetFile = "covariates.rds",
                                    url = covariates, fun = "readRDS", 
                                    destinationPath = Paths$inputPath) # For now use the ready covars
bioCov <- raster::subset(allCov, names(allCov)[1:19])
treeCov <- raster::subset(allCov, names(allCov)[20:nlayers(allCov)])
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

## For each species, make new projections
predictBiomod <- function(sp, 
                          inputPath,
                          bioCov,
                          treeCov,
                          currentTime,
                          overwritePred = FALSE){
  message(crayon::yellow(paste0("Predictions starting for ", sp, "...")))
  originalWD <- getwd()
  on.exit(setwd(originalWD))
  setwd(dirname(inputPath))
  predRas <- file.path(getwd(), paste0(sp, "/proj_future/individual_projections/", 
                                       sp,"_EMwmeanByTSS_mergedAlgo_mergedRun_mergedData.grd"))
  
  if (any(!file.exists(predRas), 
          overwritePred
          )){
    fl <- list.files(path = inputPath, 
                     pattern = "FirstModeling.models.out",
                     full.names = TRUE, 
                     recursive = TRUE)
    modInd <- get(load(file = fl))
    # Needs to decompress
    fls <- list.files(path = inputPath, pattern = "gz",
                      full.names = TRUE, 
                      recursive = TRUE)
    if (any(unique(tools::file_ext(fls)) %in% "gz")){
      lapply(fls, R.utils::gunzip)
    }
    biomod2::BIOMOD_LoadModels(modInd)
    sel.cov <- modInd@expl.var.names
    currentStack <- raster::stack(bioCov, treeCov)
    # Make sure that any missing layers are created and added
    missingLays <- setdiff(sel.cov, names(currentStack))
    if (length(missingLays) != 0){
      zeroLays <- stack(lapply(missingLays, function(lay){
        template <- currentStack[[1]] # template
        template[] <- template[]
        template[!is.na(template[])] <- 0
        names(template) <- lay
        return(template)
      })) 
    } else zeroLays <- NA
    # Now we need to select which layers will be used
    current.env.stack <- stack(currentStack[[which(names(currentStack) %in% sel.cov)]])
    # And add the missing layers
    current.env.stack <- stack(current.env.stack, zeroLays)
    # Check if all layers are there
    testthat::expect_true(all(names(current.env.stack) %in% sel.cov))
    models_proj_future <- BIOMOD_Projection(modeling.output = modInd,
                                            new.env = current.env.stack,
                                            proj.name = "future",
                                            binary.meth = "TSS", # or ROC ...
                                            output.format = ".grd",
                                            do.stack = TRUE)
    fl <- list.files(path = inputPath, 
                     pattern = "FirstModelingensemble.models.out",
                     full.names = TRUE, 
                     recursive = TRUE)
    modEns <- get(load(file = fl))
    ## Ensemble projections
    ensemble_models_proj_future <- BIOMOD_EnsembleForecasting(EM.output = modEns,
                                                              projection.output = models_proj_future,
                                                              binary.meth = "TSS", # or ROC ...
                                                              output.format = ".grd",
                                                              do.stack = FALSE)
    
    testthat::expect_true(file.exists(predRas))
    
  }
  # Now the file exists. Load, correct and return
  predRas <- raster::raster(predRas)
  # Correcting for integer forcing (x1000) we need to divide the results
  # by 1000
  predRas <- predRas/1000
  names(predRas) <- paste0(sp, "_Year", currentTime)
  message(crayon::green(paste0("Predictions done for ", 
                               sp, " for year ", currentTime)))
  return(predRas)
}

allPredictions <- lapply(1:NROW(speciesURL), function(index){
  sp <- speciesURL[index, species]
  filePath <- file.path(Paths$inputPath, sp)
  pred <- predictBiomod(sp = sp, 
                            inputPath = filePath,
                            bioCov = bioCov,
                            treeCov = treeCov,
                        currentTime = 2011)
})
names(allPredictions) <- speciesURL[["species"]]

```

